\chapter{Problem Analysis}
\label{chap:problem-analysis}

Developing a dice game that uses computer vision and artificial intelligence presents unique challenges. This chapter systematically explores these challenges and the solutions implemented in the proposed solution. It is organized into several sections, each addressing a key aspect of the project:

\begin{itemize}
    \item \textbf{Computer Vision}: Explores the technical hurdles in dice detection and recognition, including lighting variations, perspective distortions, background complexities, and real-time processing requirements. It also discusses the image preprocessing techniques employed to enhance detection accuracy.
    
    \item \textbf{Model Architecture}: Details the design and implementation of the AI model, including the selection of pre-trained object detection models, training on custom datasets, and integration with Roboflow’s Inference API for efficient dice recognition.
    
    \item \textbf{Asynchronous Processing}: Describes the use of Kotlin Coroutines to manage background tasks, ensuring a responsive user interface. This section covers coroutine scopes, launching coroutines, suspend functions, and error handling mechanisms.
    
    \item \textbf{Gameplay}: Analyzes the core game mechanics, including the scoring algorithm, adaptive AI behaviors, and the overall game flow. It highlights how these elements contribute to an engaging and balanced gameplay experience.
    
    \item \textbf{Existing Solutions}: Reviews current approaches and technologies in virtual dice games, and dice detection and recognition, comparing various models and platforms to contextualize the thesis solution's contributions and improvements.
\end{itemize}

By addressing these components, the chapter lays the foundation for understanding the technical and strategic decisions that underpin the development of a robust and engaging dice game application.

\section{Computer Vision}

Computer vision plays an important role in seamlessly integrating physical dice into digital gameplay. This section delves into the challenges and solutions associated with dice detection and recognition, highlighting the critical techniques and advancements in the field.

Computer vision is an essential component in developing applications that interact with the physical world through visual data. It encompasses a wide range of techniques and algorithms designed to enable machines to interpret and understand visual information~\cite{szeliski2010computer}. One of the foundational works in this domain is Szeliski's comprehensive guide, which explores fundamental algorithms and their applications across various sectors. This work provides invaluable insights into image processing, feature detection, and machine learning techniques that are crucial for advancing modern computer vision applications.

The advent of deep learning has revolutionized computer vision, particularly through convolutional neural networks (CNNs) that have demonstrated exceptional performance in image classification tasks~\cite{krizhevsky2012imagenet}. These networks have enabled more accurate and efficient processing of visual data, thereby enhancing the capabilities of computer vision systems.

A significant advancement in real-time object detection is the YOLOv3 architecture introduced by Redmon and Farhadi. YOLOv3 balances speed and accuracy effectively, making it highly suitable for applications requiring real-time performance~\cite{redmon2018yolov3}. This architecture has been instrumental in improving the efficiency of object detection tasks, including the recognition of dice in dynamic environments.

Current surveys underscore the continuous innovations and emerging trends in deep learning algorithms for image classification. These advancements emphasize the growing applicability and performance enhancements of computer vision technologies in various applications~\cite{khan2020survey}. The integration of these cutting-edge techniques is fundamental to overcoming the challenges inherent in dice detection and recognition, thereby facilitating a more immersive and interactive gaming experience.

\subsection{Challenges in Dice Recognition}

The implementation of accurate dice recognition presents several technical challenges:
\begin{itemize}
    \item \textbf{Lighting Variations}: Dice faces appear differently under various lighting conditions. This includes shadows that can obscure the patterns of pips on the dice, reflective surfaces causing glare, and significant differences between indoor and outdoor lighting that affect contrast and visibility.
    \item \textbf{Perspective and Orientation}: The system must handle dice captured at different angles, which affects how pips appear in the image. Multiple dice can overlap or occlude each other, and the distance between the camera and dice impacts pip visibility and overall recognition accuracy.
    \item \textbf{Background Complexity}: Various playing surfaces can affect detection reliability. Similar patterns in the background may trigger false positives, while moving backgrounds, such as when playing on unstable surfaces, can further complicate the detection process.
    \item \textbf{Real-time Processing Requirements}: The system must process frames quickly for a responsive user experience. This involves careful management of battery consumption and memory usage, requiring optimized processing algorithms and efficient resource management.
\end{itemize}

\subsection{Image Preprocessing}

The system employs a sophisticated preprocessing pipeline using Android's built-in Bitmap processing capabilities and the AndroidX Core-KTX library for image manipulation. The preprocessing steps, illustrated in Listing~\ref{lst:image_preprocess}, utilize several key components:

\begin{lstlisting}[language=Kotlin, caption={Image Preprocessing Pipeline}, label=lst:image_preprocess]
    private suspend fun preprocessImage(bitmap: Bitmap): Bitmap {
        return withContext(Dispatchers.Default) {
            try {
                // Step 1: Convert to RGB if needed
                val rgbBitmap = ensureRGBFormat(bitmap)

                // Step 2: Enhance contrast and normalize lighting
                val enhancedBitmap = enhanceContrast(rgbBitmap)

                // Step 3: Scale while maintaining aspect ratio
                val scaledBitmap = scaleWithAspectRatio(enhancedBitmap, TARGET_SIZE)

                // Step 4: Apply noise reduction
                val finalBitmap = reduceNoise(scaledBitmap)

                Timber.d("Preprocessing completed successfully")
                finalBitmap
            } catch (e: Exception) {
                Timber.e(e, "Error during image preprocessing")
                // Fallback to basic scaling if enhancement fails
                Bitmap.createScaledBitmap(bitmap, TARGET_SIZE, TARGET_SIZE, true)
            }
        }
    }
\end{lstlisting}
\begin{itemize}
    \item \textbf{RGB Conversion}: Uses Android's ColorSpace API to ensure consistent color representation
    \item \textbf{Contrast Enhancement}: Implements a histogram equalization algorithm through Android's ColorMatrix
    \item \textbf{Aspect Ratio Scaling}: Utilizes Android's createScaledBitmap with custom calculations for dimension preservation
    \item \textbf{Noise Reduction}: Applies a Gaussian blur filter using Android's RenderScript (now deprecated) or equivalent Core-KTX functions
\end{itemize}

The implementation is shown in Listing~\ref{lst:image_preprocess}, with each step carefully orchestrated to maintain image quality while optimizing for dice detection.

This preprocessing pipeline ensures optimal image quality before the detection phase, which is further detailed in the Model Architecture section.

\section{Model Architecture}

The dice recognition system leverages a pre-trained object detection model from Roboflow~\cite{bib:kavidataset}. The model processes images at 640x640 resolution and was trained on a custom dataset of 250 images, supporting six distinct classes representing dice faces 1-6. Developed and hosted on Roboflow's platform, it provides efficient object detection capabilities through their API service.

The model is then accessed through Roboflow's Hosted Inference API, with preprocessing handling as shown in Listing~\ref{lst:image_preprocess}.

The preprocessing pipeline starts by converting the image to RGB format to standardize color space, then contrasts enhancement and lighting normalization to improve visibility, particularly under varying conditions. The image is then resized while preserving its aspect ratio to maintain recognition accuracy, then noise reduction removes unwanted artifacts that could interfere with recognition, especially in suboptimal conditions. If preprocessing fails, the system defaults to basic scaling to ensure usability.

\begin{itemize}
    \item RGB format conversion: 
    \item Image scaling to the required 640x640 dimensions
    \item Confidence threshold (set at 0.4 for reliable detections)
\end{itemize} 

For each detected die, the model outputs bounding box coordinates, confidence scores, and class labels, which the application processes to update the game state.

The detection pipeline evolved throughout development, starting with a basic implementation and later expanding to include more sophisticated preprocessing and validation.
The pipeline was later enhanced to improve detection reliability through:
\begin{itemize}
    \item \textbf{Advanced Preprocessing}: Implementation of RGB format conversion, contrast enhancement, and adaptive scaling while maintaining aspect ratios.
    \item \textbf{Robust Validation}: Addition of comprehensive detection validation including aspect ratio checks, minimum size requirements, and position validation.
    \item \textbf{Quality Filters}: Implementation of confidence threshold and noise reduction techniques.
\end{itemize}

This enhanced pipeline significantly improved detection accuracy and reliability across various lighting conditions and capture scenarios.

\section{Asynchronous Processing}

Kotlin Coroutines are utilized to efficiently manage asynchronous updates, ensuring the application remains responsive. This section explores the benefits of coroutines and their role in handling background tasks within the application.

\subsection{Coroutine Scope}

The \texttt{viewModelScope} is tied to the lifecycle of the ViewModel. This ensures that coroutines are automatically canceled when the ViewModel is cleared, preventing memory leaks and unnecessary processing.

\begin{figure}[ht!]
    \centering
    \begin{tikzpicture}[box/.style={draw, rectangle, fill=gray!20, text width=3cm, align=center, minimum height=1cm}, >=Stealth]
        % Nodes
        \node[box] (start) {ViewModel Created};
        \node[box, below=2cm of start] (scope) {viewModelScope Active};
        \node[box, below=2cm of scope] (end) {ViewModel Cleared};
        % Arrows
        \draw[->] (start) -- (scope);
        \draw[->] (scope) -- (end);
    \end{tikzpicture}
\end{figure} 
\label{fig:lifecycle_viewmodelscope}

\subsection{Launching Coroutines}

The \texttt{launch} function starts a new coroutine, allowing non-blocking execution. This is crucial for processing tasks like dice value detection or data loading, which can be time-consuming. The \texttt{viewModelScope} is tied to the ViewModel's lifecycle, ensuring that any running coroutines are automatically cancelled when the ViewModel is cleared, preventing memory leaks and unnecessary background work.


\begin{lstlisting}[language=Kotlin, caption={Launching a Coroutine}, label=lst:launch_coroutine]
    viewModelScope.launch {
        // Load vibration setting
        dataStoreManager.getVibrationEnabled()
            .collect { enabled -> _vibrationEnabled.value = enabled }
    }
\end{lstlisting}

\subsection{Suspend Functions}

Suspend functions are a key feature of Kotlin's coroutine system, marking functions that can be paused and resumed. These functions can only be called from within a coroutine or another suspend function, ensuring proper asynchronous execution.
\begin{lstlisting}[language=Kotlin, caption={Suspend Function Example}, label=lst:suspend_function]
    fun rollDice() {
        if (isRolling.value || !isRollAllowed.value) return
        viewModelScope.launch {
            trackDecision()
            trackRoll()
            _isLoading.value = true
            val results = diceManager.rollDiceForBoard(_selectedBoard.value)
            if (_vibrationEnabled.value) provideHapticFeedback()
            // Process game state after rolling
            val newState = processGameState(results)
            _gameState.value = newState
        }
    }
\end{lstlisting}

In Listing~\ref{lst:suspend_function}, the \texttt{rollDice} function is designed to manage the dice-rolling process within the game. It is executed within a coroutine scope using \texttt{viewModelScope.launch}, which allows it to perform asynchronous operations without blocking the main thread. This ensures that the UI remains responsive while the dice are being rolled.

The function begins by checking if a roll is already in progress or if rolling is not allowed, returning early if either condition is true. This prevents unnecessary operations and ensures that the game logic is executed only when appropriate.

Within the coroutine, the function tracks the player's decision and roll actions, providing valuable data for game analytics. It then sets a loading state to indicate that a roll is in progress. The actual dice rolling is performed by the \texttt{diceManager}, which returns the roll results.

If vibration feedback is enabled, the function provides haptic feedback to enhance the user experience. Finally, the function processes the game state based on the roll results and updates the game state accordingly.

The use of coroutines in this function allows for efficient management of asynchronous tasks, ensuring that the game logic is executed smoothly and without interruption. This design pattern is essential for maintaining a responsive and engaging user interface in a coroutine-based architecture.

\subsection{Error Handling}

The \texttt{try-catch} block within the coroutine handles exceptions, ensuring errors are logged and managed gracefully. This prevents crashes and maintains application stability.
\begin{lstlisting}[language=Kotlin, caption={Error Handling in Coroutines}, label=lst:error_handling_coroutine]
    viewModelScope.launch {
        try {
            val detections = roboflowRepository.detectDice(bitmap)
            _detectionState.value = if (detections.isNotEmpty()) {
                DetectionState.Success(detections)
            } else {
                DetectionState.NoDetections
            }
        } catch (e: Exception) {
            Timber.e(e, "Error detecting dice")
            _detectionState.value = DetectionState.Error(e.message ?: "Unknown error")
        }
    }
\end{lstlisting}

The coroutine runs in the background, ensuring the main thread remains responsive to user interactions. Listing~\ref{lst:error_handling_coroutine} illustrates the implementation of error handling within a coroutine. It processes collected data to update a LiveData property, enabling the UI to dynamically reflect any changes.

\subsection{Interface Responsiveness}

In interactive applications, maintaining a responsive user interface (UI) is critical to delivering a seamless user experience. By offloading computationally intensive tasks, such as AI decision-making and data processing, to background threads, the main UI thread remains available for handling real-time user interactions. This approach minimizes UI lag, ensuring that animations, gestures, and updates occur smoothly without delays.

For instance, in the context of dice games, background tasks such as calculating potential AI strategies or updating game states are delegated to coroutine-based background threads in Kotlin. This concurrency model enables a separation of concerns, where the UI layer focuses solely on rendering and responding to user input while backend logic operates asynchronously. The result is a user experience that feels intuitive and highly responsive, even under computationally demanding scenarios.

\section{Gameplay}

Making a game with lively gameplay presents many challenges, particularly when integrating adaptive artificial intelligence and ensuring a balance between challenge and accessibility. This section delves into the solutions employed to address these challenges.

\subsection{Scoring Algorithm}

The scoring system implements different calculations for each game variant:

\subsubsection{Balut Scoring}
The scoring for the Balut variant is calculated as:

\begin{equation}
    S_{category} = w_c \times b_s \times (1 + (r - 1)(1 - l)) \times m_s
\end{equation}

where $S_{category}$ represents the final score for a category, $w_c$ is the category weight (1.0-2.5), $b_s$ is the base score from dice combination, $r$ is the random factor (0.5-1.5), $l$ is the AI skill level (0-1), and $m_s$ is the multiplier based on game state (0.6-1.2).

\subsubsection{Pig Scoring}
For the Pig variant, the turn score is calculated as:

\begin{equation}
    S_{turn} = \begin{cases}
        \sum_{i=1}^{n} d_i, & \text{if } d_i \neq 1 \text{ for all rolls} \\
        0, & \text{if any } d_i = 1
    \end{cases}
\end{equation}

where $S_{turn}$ represents the score for a turn, $n$ is the number of rolls in the turn, and $d_i$ is the value of the $i$th dice roll.

\subsubsection{Greed Scoring}
For the Greed variant, the turn score is calculated as:

\begin{equation}
    S_{greed} = \sum_{i=1}^{k} (v_i \times m_i) + b
\end{equation}

where $S_{greed}$ represents the final score for a turn, $k$ is the number of scoring combinations, $v_i$ is the value of each scoring combination, $m_i$ is the multiplier for each combination type (e.g., triples, straights), and $b$ is any bonus points for special combinations. The scoring combinations are:

\begin{equation}
    v_i = \begin{cases}
        1000, & \text{if triple ones} \\
        200 \times n, & \text{if triple } n \text{ (where } n > 1\text{)} \\
        100, & \text{if single one} \\
        50, & \text{if single five} \\
        1500, & \text{if straight (1-6)} \\
        750, & \text{if three pairs} \\
        0, & \text{otherwise}
    \end{cases}
\end{equation}


\subsection{Adaptive AI}
\label{subsec:adaptive_ai}

In the game, the Adaptive AI simulates a dynamic and intelligent opponent that makes strategic decisions based on the player's play style and the current game state. The AI is integrated into different parts of the gameplay, including the classic games of \emph{Pig}, \emph{Greed}, and \emph{Balut}.

In the game of \emph{Balut}, when it is the AI's turn, it must decide whether to roll, hold, or bank dice based on the current game state. This decision-making process is implemented in the \texttt{handleAITurn} function. In the game of \emph{Greed}, the AI determines which dice combinations to keep after each roll, aiming to maximize its score while minimizing risk. Meanwhile, in the game of \emph{Pig}, the AI's decision to roll or bank is influenced by both the current score and the opponent's score, as well as the overall target score. Throughout these games, a game tracker records the AI's decisions, allowing for continuous refinement of its behavior.

\subsubsection{The \texttt{handleAITurn} Function}

The \texttt{handleAITurn} function in the game of \emph{Balut} guides the AI's actions during its turn. It considers factors such as the number of rolls remaining, the values of the dice, and the potential scores of different categories. Depending on the game state, the AI either selects a category to score (when no rolls are left) or decides which dice to hold for the next roll. The Listing~\ref{lst:ai_turn_function}  illustrates the function:

\begin{lstlisting}[language=Kotlin, caption={handleAITurn Function}, label=lst:ai_turn_function]
    private fun handleAITurn(
        diceResults: List<Int>, currentState: BalutScoreState
    ): BalutScoreState {
        gameTracker.trackDecision()
        if (currentState.rollsLeft <= 0) {
            // AI chooses a category
            val category = chooseAICategory(diceResults, currentState)
            gameTracker.trackBanking(ScoreCalculator.calculateCategoryScore(diceResults, category))
            return scoreCategory(currentState, diceResults, category)
        }

        // AI decides which dice to hold
        gameTracker.trackRoll()
        val diceToHold = decideAIDiceHolds(diceResults)

        return currentState.copy(
            rollsLeft = currentState.rollsLeft - 1,
            heldDice = diceToHold
        )
    }
\end{lstlisting}

\subsubsection{The \texttt{shouldAIBank} Function: Adapting to Player Style}

In various dice games, the AI's decision-making process is influenced by the human player's observed play style. By analyzing the player's tendencies, the AI adjusts its risk-taking behavior accordingly. The implementation varies across different games but follows a common principle: \textbf{aggressive players encourage riskier AI behavior, while cautious players make the AI more conservative}.

\paragraph{Pig: Adjusting Risk in Banking Decisions}  
In the game of \emph{Pig}, the AI decides whether to bank its current turn score or continue rolling. The AI considers the player's style before making this decision. Listing~\ref{lst:player_style_adjustment} demonstrates this adjustment:

\begin{lstlisting}[language=Kotlin, caption={Player Style Adjustment}, label=lst:player_style_adjustment]
val playerAnalysis = statisticsManager.playerAnalysis.value
val playerStyle = playerAnalysis?.playStyle ?: PlayStyle.BALANCED

// Adjust based on player style
val styleAdjustment = when (playerStyle) {
    PlayStyle.AGGRESSIVE -> -0.1  // More risk-taking
    PlayStyle.CAUTIOUS -> 0.1     // More conservative
    else -> 0.0
}
\end{lstlisting}

Here, the AI retrieves the player’s style from the statistics manager. An aggressive player causes the AI to reduce its tendency to bank, encouraging riskier rolls, while a cautious player leads the AI to secure points more readily.

\paragraph{Greed: Dynamic Banking Thresholds}  
In \emph{Greed}, a more complex dice game, the AI incorporates both \textbf{a base banking threshold} and \textbf{situational adjustments}. The function in Listing~\ref{lst:greed_ai_banking} demonstrates how the AI modifies its strategy:


\begin{lstlisting}[language=Kotlin, caption={AI Banking Strategy in Greed}, label=lst:greed_ai_banking]
val baseThreshold = when (playerAnalysis?.playStyle) {
    PlayStyle.AGGRESSIVE -> 1200  // AI takes greater risks
    PlayStyle.CAUTIOUS -> 900     // AI banks earlier
    else -> 1000                  // Default threshold
}

// Adjust for game state
val situationalAdjustment = when {
    aiTotalScore == 0 -> -200  // Willing to risk when no points are banked
    aiTotalScore >= 8000 -> -300  // More aggressive near victory
    currentTurnScore >= 2000 -> 500  // Encouraged to bank large scores
    else -> 0
}

// Final threshold considers randomness for unpredictability
val finalThreshold = (baseThreshold + situationalAdjustment).coerceIn(800, 2000)
\end{lstlisting}

In this case, an aggressive opponent leads the AI to play more recklessly, increasing the threshold before banking. Conversely, a cautious opponent results in a lower threshold, making the AI secure its points earlier.

\paragraph{Balut: Play Style Influences Dice Holding and Category Selection}  
\emph{Balut} introduces additional layers of decision-making, including \textbf{dice-holding strategies} and \textbf{category selection}. The AI modifies its strategy based on the player’s style:

\textbf{Dice Holding:}  
\begin{lstlisting}[language=Kotlin, caption={AI Dice Holding in Balut}, label=lst:balut_dice_holding]
val aiStyle = when (playerAnalysis?.playStyle) {
    PlayStyle.AGGRESSIVE -> 0.7  // More likely to hold high-value dice
    PlayStyle.CAUTIOUS -> 0.5    // More selective
    else -> 0.6
}
\end{lstlisting}

Here in Listing~\ref{lst:balut_dice_holding}, an aggressive player makes the AI hold more dice in an attempt to go for high-risk, high-reward combinations.

\textbf{Category Selection:}  
\begin{lstlisting}[language=Kotlin, caption={AI Category Selection in Balut}, label=lst:balut_category_selection]
val categoryScores = availableCategories.associateWith { category ->
    val baseScore = ScoreCalculator.calculateCategoryScore(diceResults, category)
    val weight = when {
        category == "Five of a Kind" -> if (baseScore > 0) 2.5 else 0.0
        category == "Large Straight" -> if (baseScore > 0) 2.2 else 0.0
        else -> 1.0
    }
    weight * (if (playerAnalysis?.playStyle == PlayStyle.AGGRESSIVE) 1.2 else 1.0)
}
\end{lstlisting}

Here in Listing~\ref{lst:balut_category_selection}, the AI prioritizes high-value categories when facing an aggressive player, taking bigger risks. Against a cautious player, it leans towards safer choices.

Across \emph{Pig}, \emph{Greed}, and \emph{Balut}, the AI \textbf{adapts dynamically} to the player's style. This adaptability ensures a more engaging and challenging experience, making the AI feel responsive rather than static. Whether banking, holding dice, or selecting categories, the AI's decisions reflect the player's behavior, leading to a more immersive gameplay experience.


\subsection{Game Mechanics}

The game mechanics are crucial for delivering an engaging and intuitive gameplay experience. They are designed to ensure clear rules and interactions for both players and the AI, enabling a seamless game flow.
An important component of the implementation is the \texttt{handleTurn} method, as shown in Listing~\ref{lst:handleTurn}. This method differentiates between player and AI turns and manages key actions such as dice holding and roll counting. Its modular design supports clear separation of player and AI logic into distinct methods, reducing complexity and improving maintainability. This structure makes it easy to add new game modes or modify the AI behavior without disrupting existing functionality.

\begin{lstlisting}[language=Kotlin, caption={handleTurn Function}, label=lst:handleTurn]
    fun handleTurn(
            currentState: GameScoreState.PigScoreState, diceResult: Int? = null
        ): GameScoreState.PigScoreState = when (currentState.currentPlayerIndex) {
                AI_PLAYER_ID.hashCode() -> handleAITurn(currentState, diceResult)
                else -> handlePlayerTurn(currentState, diceResult)
            }
\end{lstlisting}

During its turn, the AI uses a blend of predefined rules and probabilistic decision-making to evaluate the game state and select the optimal strategy. Meanwhile, the game mechanics prioritize user experience by offering clear visual and interactive cues, ensuring that players can focus on strategy without being hindered by the interface.

In the \texttt{handleTurn} method as shown in Listing~\ref{lst:handleTurn}, the method differentiates between player and AI turns and manages key actions such as dice holding and roll counting.
\begin{itemize}
    \item \textbf{AI Turn Handling:} If the current player's index matches the AI player's identifier, the method delegates the turn to \texttt{handleAITurn}, which implements the AI's decision-making logic. This includes evaluating the game state, deciding which dice to hold, and determining whether to bank a score or re-roll.
    \item \textbf{Player Turn Handling:} If the turn belongs to a human player, the method invokes \texttt{handlePlayerTurn}, which processes the player's actions, such as selecting dice to hold and performing a roll.
\end{itemize}

By isolating player-specific and AI-specific logic into separate methods, the design enhances code readability and maintainability. This modular approach ensures that updates or adjustments to AI strategies or player interactions can be made independently, maintaining the overall flow of the game.

This structured design allows for a compelling gaming experience by promoting a dynamic AI challenge while ensuring that interactions remain clear and responsive. The thoughtful integration of adaptive AI, clear gameplay mechanics, and user-focused design ensures that the game is accessible to players of all skill levels. Additionally, the modularity of the architecture enables the seamless incorporation of advanced features, such as multiplayer modes or new game variants, without disrupting the core mechanics.

Through these technical and strategic design choices, the project delivers an engaging dice game that is robust and scalable for future enhancements.


\section{Existing Solutions}
This section provides an overview of notable solutions in the domain of dice game applications and image recognition technologies. The analysis focuses on how each solution addresses dice recognition and gaming, while also contextualizing the unique enhancements introduced by this thesis.

\subsection{D3-Deep-Dice-Detector}

The \href{https://github.com/harshmunshi/D3-Deep-Dice-Detector}{D3-Deep-Dice-Detector} utilizes deep learning techniques, specifically convolutional neural networks (CNNs), to detect and recognize dice in images. Its primary strength lies in accurately identifying dice numbers and face values under varying conditions of light and orientation~\cite{bib:D3-Dice}.

While excelling at dice recognition, the system is focused mainly on detection. The solution presented in this thesis builds upon this by incorporating real-time updates and dynamic game interactions, further enriching the user experience with multiple game variants and adaptive AI.

\subsection{Dice Scores Recognition}

\href{https://github.com/ordovas/dice-scores-recognition}{Dice Scores Recognition} automates the scoring of dice games such as Yahtzee by interpreting dice configurations captured in images~\cite{bib:Dice-Scores-Recognition}.

This solution targets predefined games with specific scoring rules. The proposed solution, however, expands its capabilities to cover multiple game variants (\emph{Pig}, \emph{Greed}, \emph{Balut}) while introducing adaptive AI for a more interactive and personalized gameplay experience.

\subsection{Zilch-Dice}

\href{https://github.com/pandulapeter/zilch-dice}{Zilch-Dice} is a Kotlin-based application designed for the Zilch dice game variant (also known as 10000), supporting Android, Linux, macOS, and Windows clients~\cite{bib:zilchdice}. It focuses on score tracking, game state management, and player interactions, all aligned with Zilch’s ruleset.

Although it caters to a single game variant, the proposed solution supports a wider array of dice games within a unified platform. This extended functionality, coupled with real-time dice detection and integrated adaptive AI, enhances player engagement and enables more dynamic gameplay.

\subsection{Flutzy}

\href{https://github.com/amuhaimin02/flutzy}{Flutzy} is a cross-platform dice game application that offers multiple game modes and an intuitive user interface~\cite{bib:flutzy}. It aims for a seamless experience across different devices.

Flutzy is still under development, whereas the proposed solution is fully realized with Android-specific optimizations, including Jetpack Compose and Clean Architecture principles. The real-time dice detection and interactive gameplay in the proposed system provide a more responsive experience compared to Flutzy’s current static modes.

\subsection{Python-Dice}

\href{https://github.com/jckuhl/Python-Dice}{Python-Dice} contains Python scripts for simulating dice rolls and calculating scores according to various game rules~\cite{bib:python-dice}. It serves as a backend tool for dice game logic without a dedicated user interface or real-time capabilities.

The proposed solution goes beyond simulation and scoring by offering a complete Android application with a user-friendly interface and integrated real-time dice detection, making it more suitable for end-users looking for an engaging, interactive dice gaming experience.

\subsection{Dice Detection}

\href{https://github.com/binaryshrey/Dice}{Dice Detection} by Nell Byler specializes in detecting and counting dice within images using image processing techniques~\cite{bib:nell-byler}. It focuses on accurately identifying dice face values and quantities.

This thesis extends the capabilities of existing implementations with several key innovations:

\begin{itemize}
    \item \textbf{Unified Approach}: Seamlessly integrates real-time dice detection with multiple game variants in a single application.
    
    \item \textbf{Adaptive AI}: Implements an intelligent opponent that learns from player behavior, unlike existing solutions that use static game logic.
    
    \item \textbf{Modern Architecture}: Utilizes current Android development practices with Jetpack Compose and Clean Architecture, providing better maintainability and user experience.
    
    \item \textbf{Real-Time Processing}: Offers immediate dice recognition and game state updates, compared to the static or delayed processing in existing solutions.
\end{itemize}
